---
title: "Final Report: Election Prediction"
author: "Alexander Bendeck, Lynn Fan, Cathy Lee, Alice Liao, Justina Zou"
date: "28 October 2020"
fontsize: 11pt
geometry: "left=1.5cm,right=1.5cm,top=1.5cm,bottom=1.5cm"
output: 
  pdf_document: 
  latex_engine: xelatex
  fig_width: 2
  fig_height: 1
header-includes:
- \usepackage{booktabs}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pkgtest, echo=FALSE}
# source: https://stackoverflow.com/questions/9341635/check-for-installed-packages-before-running-install-packages
pkgTest <- function(x)
  {
    if (!require(x,character.only = TRUE))
    {
      install.packages(x,repos = "http://cran.r-project.org",dep=TRUE)
        if(!require(x,character.only = TRUE)) stop("Package not found")
    }
  }
```

```{r check package,include=FALSE}
pkgTest("gtable") 
pkgTest("ggplot2") 
pkgTest("gridExtra") 
pkgTest("xtable") 
pkgTest("coda") 
pkgTest("dplyr")
pkgTest("reshape2")
pkgTest("readr")
pkgTest("grid")
pkgTest("bayesplot")
pkgTest("knitr")
pkgTest("tidyverse")
pkgTest("stringr")
pkgTest("stringi")
pkgTest("ggplotify")
pkgTest("tidybayes") # easy residuals
```

```{r packages, include=FALSE, message = F}
library(gtable)
library(ggplot2)
library(gridExtra)
library(rjags)
library(R2jags)
library(grid)
library(xtable)
library(coda)
library(bayesplot)
library(rstanarm)
library(dplyr)
library(reshape2)
library(knitr)
library(tidyverse)
library(stringr)
library(stringi)
library(ggplotify)
library(brms)
library(tidybayes) # easy residuals
library(usmap)
#library(xtable)
```

## Introduction

## Data Description

Senate poll/House poll: from https://data.fivethirtyeight.com/
States_cov_matrix: from the Economist paper
Economist prediction polls: from the Economist paper (pres)
partisan lean (prior for pres): from 538
Abramowitz data: from the Economist paper (pres)


## Exploratory Data Analysis

```{r load data}
pres_eda_df = readRDS("cleaned_pres_polls.rds")
econ_eda_df = readRDS("cleaned_econ_polls.rds")
econ_state_df = econ_eda_df %>%
  group_by(state) %>%
  summarise(n=n(),avg=mean(y))
house_eda_df = readRDS("cleaned_house_polls.rds")
senate_eda_df = readRDS("cleaned_senate_polls.rds")
senate_state_df = senate_eda_df %>%
  group_by(state) %>%
  summarise(n=n(),avg=mean(y))
```

```{r president EDA}
#number of polls for each state in the economist paper data set
plot_usmap(data = econ_state_df, values = "n", color = "red") + 
  scale_fill_continuous(name = "Number of Polls", label = scales::comma) + 
  theme(legend.position = "right")
#democrat percentage in polling data
ggplot(econ_state_df, aes(days_to_election, y)) + geom_bin2d()
#democrat percentage by states
plot_usmap(data = econ_state_df, values = "avg", color = "black") + 
  scale_fill_continuous(low = "red", high = "white",name = "Average Democrat Vote Share", label = scales::comma) + 
  theme(legend.position = "right")
```

```{r senate EDA}
#number of polls for senate
plot_usmap(data = senate_state_df, values = "n", color = "red") + 
  scale_fill_continuous(name = "Number of Polls", label = scales::comma) + 
  theme(legend.position = "right")
#biden percentage in polling data
ggplot(senate_eda_df, aes(days_to_election, y)) + geom_bin2d()
#democrat percentage by states
plot_usmap(data = senate_state_df, values = "avg", color = "black") + 
  scale_fill_continuous(low = "red", high = "white",name = "Average Democrat Vote Share", label = scales::comma) + 
  theme(legend.position = "right")
```
```{r house EDA}
#biden percentage in polling data
ggplot(house_eda_df, aes(x=days_to_election, y=y,color=as.factor(seat_number))) + geom_point()
```

## Method

## Results

## Model Validation and Sensitivity Checks

## Diagnostics

# Appendix

## References
1. Linzer, D. A. (2013). Dynamic Bayesian Forecasting of Presidential Elections in the States. Journal of the American Statistical Association, 108(501), 124-134. doi:10.1080/01621459.2012.737735
\newline
2. Hansford, T. G., &amp; Gomez, B. T. (2010). Estimating the Electoral Effects of Voter Turnout. American Political Science Review, 104(2), 268-288. doi:10.1017/s0003055410000109
\newline
3. 2020 Election. (2020, October 20). Retrieved October 20, 2020, from https://fivethirtyeight.com/politics/elections/
\newline
4. Park, D. K., Gelman, A., &amp; Bafumi, J. (2006). State-Level Opinions from National Surveys:. Public Opinion in State Politics, 209-228. doi:10.2307/j.ctvr33bdg.17
\newline
5. Mahler, V. A., Jesuit, D. K., &amp; Paradowski, P. R. (2013). Electoral Turnout and State Redistribution. Political Research Quarterly, 67(2), 361-373. doi:10.1177/1065912913509306
\newline
6. Uhlaner, C. J., &amp; Scola, B. (2015). Collective Representation as a Mobilizer. State Politics &amp; Policy Quarterly, 16(2), 227-263. doi:10.1177/1532440015603576
\newline
7. Godbout, J. (2012). Turnout and presidential coattails in congressional elections. Public Choice, 157(1-2), 333-356. doi:10.1007/s11127-012-9947-7
\newline
8. Kim, S. S., Alvarez, R. M., &amp; Ramirez, C. M. (2020). Who Voted in 2016? Using Fuzzy Forests to Understand Voter Turnout. doi:10.33774/apsa-2020-xzx29
\newline
9. Weinschenk, A. C. (2019) That’s Why the Lady Lost to the Trump: Demographics and the 2016 Presidential Election, Journal of Political Marketing, 18:1-2, 69-91, DOI: 10.1080/15377857.2018.1478657
\newline
10. Charles, K. K., &amp; Stephens, M. (2011). Employment, Wages and Voter Turnout. doi:10.3386/w17270
\newline
11. Hills, M. (2020, September 25). US election 2020: A really simple guide. Retrieved October 20, 2020, from https://www.bbc.com/news/election-us-2020-53785985
\newline
12. Railey, K. (2016). Federal Judges Let Stand North Carolina,'s New Congressional Map. The Hotline. https://link.gale.com/apps/doc/A498010836/ITOF?u=duke_perkins&sid=ITOF&xid=119d6ad9
\newline
13. Perrin, A. J., &amp; Ifatunji, M. A. (2020). Race, Immigration, and Support for Donald Trump: Evidence From the 2018 North Carolina Election. Sociological Forum, 35(S1), 941-953. doi:10.1111/socf.12600
\newline
14. Redistricting in North Carolina. (2020). Retrieved October 21, 2020, from https://ballotpedia.org/Redistricting_in_North_Carolina
\newline
15. Â§ 132-1. Public Records.
https://www.ncleg.gov/EnactedLegislation/Statutes/PDF/BySection/Chapter_132/GS_132-1.pdf
\newline
16. Â§ 163-82.10.  Official record of voter registration.
https://www.ncleg.gov/EnactedLegislation/Statutes/PDF/BySection/Chapter_163/GS_163-82.10.pdf
\newline
17. FairVote.org. (n.d.). Voter Turnout. Retrieved October 22, 2020, from https://www.fairvote.org/voter_turnout
\newline
18. McDonald, M. P. (2020). Voter Turnout Demographics. Retrieved October 23, 2020, from http://www.electproject.org/home/voter-turnout/demographics
\newline
19. Y. Ghitza, A. Gelman’s (2013). Deep interactions with MRP: Election Turnout and Voting Patterns Among Small Electoral Subgroups. American Journal of Political Science 57 762– 776.
\newline
20. Walker, H.L., Herron, M.C. & Smith, D.A. Early Voting Changes and Voter Turnout: North Carolina in the 2016 General Election. Polit Behav 41, 841–869 (2019). https://doi-org.proxy.lib.duke.edu/10.1007/s11109-018-9473-5 
\newline
21. Pew Research Center (2020). Men and women in the U.S. continue to differ in voter turnout rate, party identification. https://www.pewresearch.org/fact-tank/2020/08/18/men-and-women-in-the-u-s-continue-to-differ-in-voter-turnout-rate-party-identification/ 
\newline
22. Pew Research Center (2016). Party affiliation among voters: 1992-2016.
https://www.pewresearch.org/politics/2016/09/13/2-party-affiliation-among-voters-1992-2016/ 
\newline
23. J. Misra (2019). Voter Turnout Rates Among All Voting Age and Major Racial and Ethnic Groups Were Higher Than in 2014. https://www.pewresearch.org/fact-tank/2020/08/18/men-and-women-in-the-u-s-continue-to-differ-in-voter-turnout-rate-party-identification/
\newline
24. Killian, L. (2012). Five Myths About Independent Voters. Retrieved October 22, 2020, from https://www.washingtonpost.com/opinions/five-myths-about-independent-voters/2012/05/17/gIQAZmGyWU_story.html
\newline
25. Igielnik, R. (2020). Men and Women in the U.S. Continue to Differ in Voter Turnout Rate. Retrieved October 22, 2020, from https://www.pewresearch.org/fact-tank/2020/08/18/men-and-women-in-the-u-s-continue-to-differ-in-voter-turnout-rate-party-identification/
\newline
26. Akee, R. (2019). Voting and Income. Retrieved October 22, 2020, from https://econofact.org/voting-and-income

