---
title: "Interim Report: Who Votes in NC?"
author: "Alexander Bendeck, Lynn Fan, Cathy Lee, Alice Liao, Justina Zou"
date: "20 October 2020"
fontsize: 11pt
output: 
  pdf_document: 
  latex_engine: xelatex
  fig_width: 2
  fig_height: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(gtable)
library(ggplot2)
library(gridExtra)
library(rjags)
library(R2jags)
library(grid)
library(xtable)
library(coda)
library(bayesplot)
library(rstanarm)
library(dplyr)
library(reshape2)
library(knitr)
library(tidyverse)
library(stringr)
library(stringi)
```

# Introduction

The United States is getting closer to the 2020 Congressional Elections on November 3, 2020. With the current polarizing political landscape, the congressional election outcomes are significant to determine the next stage of this country. As it becomes critical for statisticians to help build forecasting models to predict the election outcomes, we need to first understand the patterns of voter turnout. Voting has special importance in democratic systems, but only about half of the eligible U.S. citizenry votes, and there are real political consequences when voters differ systematically from nonvoters (Uhlaner et al.). There has been abundant literature proving that variation in voter turnout will have electoral consequences (Hansford, et al.), in a number of ways. First of all, the media conventional wisdom is that “higher turnout benefits Democrats,” although there has been mixed evidence about this theory (Weinschenk, 2019). Second, literature has proven certain demographic factors to statistically significantly benefit one party over the other, in both congressional elections and presidential races. For example, gender, race and party registration could help profile voting patterns for congressional elections (Uhlaner et al.). Election prediction models need the baseline population of voters to predict the potential outcomes, and the demographic composition of voters will directly determine the forecasting results. 
\newline
Among all states, North Carolina has been as a swing state in presidential and congressional elections for decades. In 2008, Obama won the state narrowly, but lost it narrowly after 4 years in 2012. Since 1996, the Republican statewide vote share in congressional elections has varied "from a low of 45% in 2008 to a high of 55% in 2014 (Perrin et al.)." It makes North Carolina an interesting battleground in which voter demographic changes could potentially lead to significant implications of election outcomes and "an excellent site for those interested in partisan voting trends (Perrin et al.)." This report seeks to understand the voter turnout of North Carolina for 2020 NC Congressional Elections, predicting who will vote in 2020.

# Data Description

We are using public data provided by the NC State Board of Elections, which can all be accessed directly at the link https://dl.ncsbe.gov/list.html. The database contains voter history information for elections within the past 10 years in the ncvhis files, and all legally available voter specific information ion the ncvoter files. The ncvoter files contain point-in-time snapshot voter registration data. For privacy concerns, names, birth dates and drivers license are not included, but the two types of files could be matched by North Carolina identification (NCID) number. The database was last updated on September 9, 2020. While we understand that voters might register later than that as the voter registration deadline for North Carolina is October 9, 2020, we believe it is sufficient to represent the majority of NC potential voters.
\newline
From the ncvhis files, we only kept the voters that voted for the 2016 general election for our analysis. Studies have shown that presidential elections help mobilize voters, so voter turnout in presidential election years are significantly higher. In recent elections, voter turnout during presidential election years is around 60%, and only about 40% during midterm elections (FairVote.org). For North Carolina, voter turnout data in 2018 is also inappropriate to use because neither of North Carolina’s U.S. senators nor the governor was up for reelection, further demotivated voters (Perrin et al.). From the ncvoter files, we filtered demographic factors that are supported by existing literature to be significant in understanding voting patterns, including gender, race, party registration, and age (Kim et al.). We also have their county and congressional district information available. 
\newline
Additionally, we found relevant literature proving the relationship between voter turnout and wage (Charles et al.), so we found county-level median household income data from Economic Research Service under United States Department of Agriculture (https://www.ers.usda.gov/data-products/county-level-data-sets)

## Missing Data and Data Cleanup



# Method


# Appendix
## References

1. Linzer, D. A. (2013). Dynamic Bayesian Forecasting of Presidential Elections in the States. Journal of the American Statistical Association, 108(501), 124-134. doi:10.1080/01621459.2012.737735
\newline
2. Hansford, T. G., &amp; Gomez, B. T. (2010). Estimating the Electoral Effects of Voter Turnout. American Political Science Review, 104(2), 268-288. doi:10.1017/s0003055410000109
\newline
3. 2020 Election. (2020, October 20). Retrieved October 20, 2020, from https://fivethirtyeight.com/politics/elections/
\newline
4. Park, D. K., Gelman, A., &amp; Bafumi, J. (2006). State-Level Opinions from National Surveys:. Public Opinion in State Politics, 209-228. doi:10.2307/j.ctvr33bdg.17
\newline
5. Mahler, V. A., Jesuit, D. K., &amp; Paradowski, P. R. (2013). Electoral Turnout and State Redistribution. Political Research Quarterly, 67(2), 361-373. doi:10.1177/1065912913509306
\newline
6. Uhlaner, C. J., &amp; Scola, B. (2015). Collective Representation as a Mobilizer. State Politics &amp; Policy Quarterly, 16(2), 227-263. doi:10.1177/1532440015603576
\newline
7. Godbout, J. (2012). Turnout and presidential coattails in congressional elections. Public Choice, 157(1-2), 333-356. doi:10.1007/s11127-012-9947-7
\newline
8. Kim, S. S., Alvarez, R. M., &amp; Ramirez, C. M. (2020). Who Voted in 2016? Using Fuzzy Forests to Understand Voter Turnout. doi:10.33774/apsa-2020-xzx29
\newline
9. Weinschenk, A. C. (2019) That’s Why the Lady Lost to the Trump: Demographics and the 2016 Presidential Election, Journal of Political Marketing, 18:1-2, 69-91, DOI: 10.1080/15377857.2018.1478657
\newline
10. Charles, K. K., &amp; Stephens, M. (2011). Employment, Wages and Voter Turnout. doi:10.3386/w17270
\newline
11. Hills, M. (2020, September 25). US election 2020: A really simple guide. Retrieved October 20, 2020, from https://www.bbc.com/news/election-us-2020-53785985
\newline
12. Railey, K. (2016). Federal Judges Let Stand North Carolina,'s New Congressional Map. The Hotline. https://link.gale.com/apps/doc/A498010836/ITOF?u=duke_perkins&sid=ITOF&xid=119d6ad9
\newline
13. Perrin, A. J., &amp; Ifatunji, M. A. (2020). Race, Immigration, and Support for Donald Trump: Evidence From the 2018 North Carolina Election. Sociological Forum, 35(S1), 941-953. doi:10.1111/socf.12600
\newline
14. Redistricting in North Carolina. (2020). Retrieved October 21, 2020, from https://ballotpedia.org/Redistricting_in_North_Carolina
\newline
15. Â§ 132-1. Public Records.
https://www.ncleg.gov/EnactedLegislation/Statutes/PDF/BySection/Chapter_132/GS_132-1.pdf
\newline
16. Â§ 163-82.10.  Official record of voter registration.
https://www.ncleg.gov/EnactedLegislation/Statutes/PDF/BySection/Chapter_163/GS_163-82.10.pdf
\newline
17. FairVote.org. (n.d.). Voter Turnout. Retrieved October 22, 2020, from https://www.fairvote.org/voter_turnout
