---
title: "Interim Report: Who Votes in NC?"
author: "Alexander Bendeck, Lynn Fan, Cathy Lee, Alice Liao, Justina Zou"
date: "20 October 2020"
fontsize: 11pt
output: 
  pdf_document: 
  latex_engine: xelatex
  fig_width: 2
  fig_height: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(gtable)
library(ggplot2)
library(gridExtra)
library(rjags)
library(R2jags)
library(grid)
library(xtable)
library(coda)
library(bayesplot)
library(rstanarm)
library(dplyr)
library(reshape2)
library(knitr)
library(tidyverse)
library(stringr)
library(stringi)
```

# Introduction

The United States is getting closer to the 2020 Congressional Elections on November 3, 2020. With the current polarizing political landscape, the congressional election outcomes are significant to determine the next stage of this country. As it becomes critical for statisticians to help build forecasting models to predict the election outcomes, we need to first understand the patterns of voter turnout. Voting has special importance in democratic systems, but only about half of the eligible U.S. citizenry votes, and there are real political consequences when voters differ systematically from nonvoters (Uhlaner et al.). There has been abundant literature proving that variation in voter turnout will have electoral consequences (Hansford, et al.), in a number of ways. First of all, the media conventional wisdom is that “higher turnout benefits Democrats,” although there has been mixed evidence about this theory (Weinschenk, 2019). Second, literature has proven certain demographic factors to statistically significantly benefit one party over the other, in both congressional elections and presidential races. For example, gender, race and party registration could help profile voting patterns for congressional elections (Uhlaner et al.). Election prediction models need the baseline population of voters to predict the potential outcomes, and the demographic composition of voters will directly determine the forecasting results. 
\newline
Among all states, North Carolina has been as a swing state in presidential and congressional elections for decades. In 2008, Obama won the state narrowly, but lost it narrowly after 4 years in 2012. Since 1996, the Republican statewide vote share in congressional elections has varied "from a low of 45% in 2008 to a high of 55% in 2014 (Perrin et al.)." It makes North Carolina an interesting battleground in which voter demographic changes could potentially lead to significant implications of election outcomes and "an excellent site for those interested in partisan voting trends (Perrin et al.)." This report seeks to understand the voter turnout of North Carolina for 2020 NC Congressional Elections, predicting who will vote in 2020.  

# Data Description

We are using public data provided by the NC State Board of Elections, which can all be accessed directly at the link https://dl.ncsbe.gov/list.html. The database contains voter history information for elections within the past 10 years in the ncvhis files, and all legally available voter specific information ion the ncvoter files. The ncvoter files contain point-in-time snapshot voter registration data. For privacy concerns, names, birth dates and drivers license are not included, but the two types of files could be matched by North Carolina identification (NCID) number. The database was last updated on September 9, 2020. While we understand that voters might register later than that as the voter registration deadline for North Carolina is October 9, 2020, we believe it is sufficient to represent the majority of NC potential voters.  



Why we selected only the year of 2016? Significance of presidential election year, but needs literature review to confirm.

Things on 2018 in Perrin

While tracking political opinion in purple states is important, the 2018 midterm election was unusual in that there were no statewide races (what local pundits call a “blue moon” election). That is, neither of North Carolina’s U.S. senators nor the governor was up for reelection, so all races on the ballot were relatively local; the largest districts were those for U.S. Congress. Democrats won a narrow victory (51.6%) in terms of the total number of votes received in contested House races, but due to gerrymandered districts, only 3 of the state’s 13 seats remained Democratic. In addition, there were six ballot measures supported by the state’s Republican legislative majority; of these, two were rejected, but four were approved.
The 2018 midterm elections also took place at an “unsettled time” (Swidler 1986) in North Carolina (and American) history. The partisan divide in the state had raised bitter debates over gerrymandering, transgender access to public restrooms, funding and governance of public universities, teacher pay, and many others. Understanding the dynamics of identity and voting decisions in 2018 pro- vides a window into the ways these decisions may respond to different unsettled environments as well. As a swing state with demographic shifts mirroring many national trends, North Carolina is an excellent site for detailed political study.

## Data Cleanup and Variable Transformation

# Method

## Variable Selection

Why we selected gender, race, party to be significant, among others (from Uhlaner):
\newline
We focus on race, ethnicity, and gender as they are important categories in contemporary U.S. politics. Our study thus adds to the minority empowerment literature. But treating these categories separately misses a crucial point: they intersect, and that intersectionality matters. White women have different experiences than Latinas; African American men have different experiences than African American women. In this article, we assess whether being collectively descriptively represented affects turnout by looking at seven elections, the presidential races of 2000, 2004, 2008, and 2012, and the off-year elections of 2002, 2006, and 2010 to assess the intersectional impact of collective descriptive representation on voter participation.
There are well-established results on the participation of people of different gender, ethnicity, and race. Overall, historically white males had the highest voting rates in the United States, but the gender edge evaporated in the 1980s while racial and ethnic dif- ferences have narrowed, but remain. The turnout rates for men and women became essentially equal in 1980 and 1982, and ever since they have been higher for women (Center for American Women and Politics 2011).4 Note, however, that for most politi- cal activities other than voting, men still outpace women (Conway, Steuernagel, and Ahern 2005, 94–97; Verba, Schlozman, and Brady 1995, 254–57). African Americans also, historically, participated less than other Americans. That gap has been decreas- ing, reflecting the Civil Rights movement and other changes, until by 1990 the African American turnout rate was close to that of whites. Obama’s candidacy and presidency has further mobilized African American voting; in 2008, turnout of African Americans was just 1.4 points below that of non-Hispanic whites and moved 2 points higher in 2012 (File 2013, 3).5 Latinos vote at lower rates than either African Americans or whites, even when the comparison is, appropriately, limited to citizens. 
\newline
Charles on employment, wage and voter turnout: why we selected median income to be significant
\newline
Using county-level data that we have assembled for elections between 1969 and 2000, and employing a variety of OLS and TSLS models, we find that increases in local per capita earnings and log employment per adult reduce voter turnout in gubernatorial, congressional, senatorial and state legislative elections, but have no statistically significant effect on turnout in presidential contests. 

# Appendix
## References

1. Linzer, D. A. (2013). Dynamic Bayesian Forecasting of Presidential Elections in the States. Journal of the American Statistical Association, 108(501), 124-134. doi:10.1080/01621459.2012.737735
\newline
2. Hansford, T. G., &amp; Gomez, B. T. (2010). Estimating the Electoral Effects of Voter Turnout. American Political Science Review, 104(2), 268-288. doi:10.1017/s0003055410000109
\newline
3. 2020 Election. (2020, October 20). Retrieved October 20, 2020, from https://fivethirtyeight.com/politics/elections/
\newline
4. Park, D. K., Gelman, A., &amp; Bafumi, J. (2006). State-Level Opinions from National Surveys:. Public Opinion in State Politics, 209-228. doi:10.2307/j.ctvr33bdg.17
\newline
5. Mahler, V. A., Jesuit, D. K., &amp; Paradowski, P. R. (2013). Electoral Turnout and State Redistribution. Political Research Quarterly, 67(2), 361-373. doi:10.1177/1065912913509306
\newline
6. Uhlaner, C. J., &amp; Scola, B. (2015). Collective Representation as a Mobilizer. State Politics &amp; Policy Quarterly, 16(2), 227-263. doi:10.1177/1532440015603576
\newline
7. Godbout, J. (2012). Turnout and presidential coattails in congressional elections. Public Choice, 157(1-2), 333-356. doi:10.1007/s11127-012-9947-7
\newline
8. Kim, S. S., Alvarez, R. M., &amp; Ramirez, C. M. (2020). Who Voted in 2016? Using Fuzzy Forests to Understand Voter Turnout. doi:10.33774/apsa-2020-xzx29
\newline
9. Weinschenk, A. C. (2019) That’s Why the Lady Lost to the Trump: Demographics and the 2016 Presidential Election, Journal of Political Marketing, 18:1-2, 69-91, DOI: 10.1080/15377857.2018.1478657
\newline
10. Charles, K. K., &amp; Stephens, M. (2011). Employment, Wages and Voter Turnout. doi:10.3386/w17270
\newline
11. Hills, M. (2020, September 25). US election 2020: A really simple guide. Retrieved October 20, 2020, from https://www.bbc.com/news/election-us-2020-53785985
\newline
12. Railey, K. (2016). Federal Judges Let Stand North Carolina,'s New Congressional Map. The Hotline. https://link.gale.com/apps/doc/A498010836/ITOF?u=duke_perkins&sid=ITOF&xid=119d6ad9
\newline
13. Perrin, A. J., &amp; Ifatunji, M. A. (2020). Race, Immigration, and Support for Donald Trump: Evidence From the 2018 North Carolina Election. Sociological Forum, 35(S1), 941-953. doi:10.1111/socf.12600
\newline
14. Redistricting in North Carolina. (2020). Retrieved October 21, 2020, from https://ballotpedia.org/Redistricting_in_North_Carolina
\newline
15. Â§ 132-1. Public Records.
https://www.ncleg.gov/EnactedLegislation/Statutes/PDF/BySection/Chapter_132/GS_132-1.pdf
\newline
16. Â§ 163-82.10.  Official record of voter registration.
https://www.ncleg.gov/EnactedLegislation/Statutes/PDF/BySection/Chapter_163/GS_163-82.10.pdf
